---
title: "01-dRy-importing-and-functions"
format: html
---

this is ella changing something

# Don't Repeat Yourself in R
By Ella Barnes and Karina Kumar - Data Fellows of the Media Innovation Group at University of Texas at Austin.

## Goals of this session
- Import multiple files at once
- Create your own function
- Generate parameterized reports quickly

Why dRy? Our coding often requires repetitive processes. These tools can help you unnecessary repetition and maximize your efficiency. 

## The data

## How this lesson works

# Importing multiple files at once

## Set up

Importing our libraries

```{r}

#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(janitor)

```

Creating a list of file names

```{r}
#| label: import

taylor_files_list <- list.files( # <1>
  "data-processed-taylor", # <2>
  pattern = ".csv", # <3>
  full.names = T   #<4>
)

taylor_files_list
```

1. list.files()
2. where to get the file
3. pattern = how to find the files
4. full.names = T gets the full path name rather than just the file name

```{r}
#| label: combine-files 

# add set_names(basename) to create a column that has the file name
# add message and warning false 

taylor_songs <- taylor_files_list |> map(
  read_csv,
  col_types = cols(album = col_character())   # comment this out to see what happens 
) |> list_rbind() |> 
  clean_names()

taylor_songs |> glimpse()
```

1. map
2. read_csv
3. col_types
4. list_rbind()
5. clean_names()

# OYO: Importing multiple files at once - Power Outages

```{r}
#| label: import-oyo

#data-processed-power

power_files_list <- 

```

```{r}
#| label: combine-files-oyo

#power_outages <-  NA

#power_outages |> glimpse()
```

How did we get to the chart...? 

# Creating your own function - Do a chart

```{r}

```


# OYO: Creating your own function

```{r}

```


