---
title: "Taycleaning"
---

## Initial Questions

- Which album/song has the most acousticnesss?...for all categories
- Which song is the longest? Shortest?
- Which release year had the highest popularity?
- Are longer or shorter songs more popular?
- Tempo vs danceability?


## Set up

```{r}

#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(janitor)

```

```{r}
#| label: import

taylor_files_list <- list.files(
  "data-processed-taylor",
  pattern = ".csv",
  full.names = T
)

taylor_files_list
```

```{r}
#| label: combine-files 

taylor_all <- taylor_files_list |> map(
  read_csv,
  col_types = cols(album = col_character())
) |> list_rbind() |> 
  clean_names()

taylor_all |> glimpse()
```

## Which album/song has the most acousticness?...for all categories

```{r}

acousticness <- taylor_all |> group_by (album) |> 
  summarize(avg_accoustic = mean(acousticness, na.rm = T) |> round(2)) |> 
  arrange(avg_accoustic |> desc())


acousticness
```

```{r}
danceability <- taylor_all |> group_by (album) |> 
  summarize(avg_danceability = mean(danceability, na.rm = T) |> round(2)) |> 
  arrange(avg_danceability |> desc())


danceability
```

```{r}
instrumentalism <- taylor_all |> group_by (album) |> 
  summarize(avg_instrumentalism = mean(instrumentalness, na.rm = T) |> round(2)) |> 
  arrange(avg_instrumentalism |> desc())


instrumentalism
```

```{r}
energy <- taylor_all |> group_by (album) |> 
  summarize(avg_energy = mean(energy, na.rm = T) |> round(2)) |> 
  arrange(avg_energy |> desc())


energy
```

```{r}
tempo <- taylor_all |> group_by (album) |> 
  summarize(avg_tempo = mean(tempo, na.rm = T) |> round(2)) |> 
  arrange(avg_tempo |> desc())


tempo
```


```{r}
popularity <- taylor_all |> group_by (album) |> 
  summarize(avg_popularity = mean(popularity, na.rm = T) |> round(2)) |> 
  arrange(avg_popularity |> desc())


popularity
```

## Which song is the longest? Shortest?

```{r}
shortest_longest <- taylor_all |> group_by(album) |> 
  summarize(shortest_song = name[which.min(duration_min)],
            longest_song = name[which.max(duration_min)]) 

shortest_longest

```

## Which release year had the highest popularity?

```{r}
year_popularity <- taylor_all |> group_by (year(release_date)) |> 
  summarize(avg_popularity = mean(popularity),
            song_count = n()) |> 
  arrange(avg_popularity |> desc())

year_popularity
```

## Are shorter or longer songs more popular?

```{r}
ggplot(taylor_all, aes(x = duration_min, y = popularity)) +
  geom_point()
```

## Tempo vs popularity

```{r}
ggplot(taylor_all, aes(x = tempo, y = popularity)) +
  geom_point()
```

## Tempo vs danceability

```{r}
ggplot(taylor_all, aes(x = tempo, y = danceability)) +
  geom_point()
```

## Energy vs danceability

```{r}
ggplot(taylor_all, aes(x = energy, y = danceability)) +
  geom_point()
```


## Energy vs popularity

```{r}
ggplot(taylor_all, aes(x = energy, y = popularity)) +
  geom_point()
```


## Danceability vs popularity

```{r}
ggplot(taylor_all, aes(x = danceability, y = popularity)) +
  geom_point()
```

